Seshadri constants of curve configurations on surfaces

### 2024-03-21 Response/Response to reviewer.tex

\numberwithin{equation}{section}
\newtheorem{theorem}{Theorem}[section]
\newtheorem*{theorem*}{Theorem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{definition}
\newtheorem{question}[theorem]{Question}

\begin{document}
\title{Response to Reviewer}
\begin{description}
\item[Paper ID]  GMJ 231005-Subramaniam

\item [Title] Seshadri Constants of Curve Configurations on Surfaces
\item [Authors] Krishna Hanumanthu, Praveen Roy and Aditya Subramaniam
\end{description}

We thank the referee for carefully reading the manuscript and suggesting 
many useful changes including a valuable suggestion to remove the hypothesis that all the curves in $\mathcal{C}$ are linearly equivalent to a fixed divisor. We were able to do this and our main theorem now (Theorem 2.2 in the new version) is a more general result without the old assumption.

@@

We also thank the referee for pointing out an error in  Assumption 2.1 which leads to 
ambiguities at several places. We fixed this error by assuming that the arrangement is connected.

We  agree with the referee that Example 2.22 (in the new version)  can be shortened and it is better to  
focus solely on the computation of Seshadri constants.

All these suggestions substantially improved the paper and we are grateful to the referee for them.

### 2024-03-21 sc-curve-config.tex
Final-after-1st-revision-1

Finally, we give an example (Example \ref{K3}) of a computation of multi-point the Seshadri constant at the singular locus of an arrangement of curves on a K3 surface. In this example, the nefness hypothesis of Theorem \ref{main-theorem} is not true but the conclusion about the Seshadri constant still holds.

@@

  For each $1\le i \le d$, let $b_i =  \# \, C_{i} \cap {\rm Sing}(\mathcal{C})$. Since our arrangement is connected, $b_i >0$ for all $i$.
   We define the \emph{base constant} of $\mathcal{C}$ as 
   $$\text{bs}(\mathcal{C}) = \max\{b_1,\dots, b_d\},$$ i.e., 
$\text{bs}(\mathcal{C})$		
  is equal to the maximum number of singular points  of $\mathcal{C}$ that are contained in a single curve $C_{i} \in \mathcal{C}$.
	%\end{definition}

@@

By \cite[Example 3.5]{JP}, the above question has a negative answer for the Hesse arrangement of conics 
in $\mathbb{P}^2$. However it is open for line arrangements in $\mathbb{P}^2$. 
It is interesting to study the situations in which it is true on arbitrary surfaces. Our main theorem below answers this question positively under an additional nefness condition.

@@

	 Suppose that there exists an irreducible and reduced curve $D\notin \mathcal{C}$ having multiplicity $m_{p}(D)$ at each point $p \in \mathcal{P}$ such that $D \cap \mathcal{P} \ne \emptyset$ and 
		%$$\frac{L \cdot D}{\sum_{p \in \mathcal{P}} m_{p}(D)} < \frac{L\cdot C_1}{C_1^2(d-1)}.$$
		$$\frac{L \cdot D}{\sum_{p \in \mathcal{P}} m_{p}(D)} < \min_{1\le i \le d}\left\{\frac{L\cdot C_i}{b_i}\right\}.$$
So 
                $$b_i(L \cdot D)  < (L\cdot C_i) \sum_{p \in \mathcal{P}} m_{p}(D).$$ for each $i$. 
Adding over all $i$, we have 
		$$(\triangle) : \quad (L \cdot D)\sum b_i < \left(\sum_{p \in \mathcal{P}} m_{p}(D)\right)\sum L\cdot C_i.$$
		By B\'ezout's theorem applied to $D$ and $C_1+\dots+C_d$, we obtain
		%\begin{eqnarray*}
		%D\cdot C_i &\geq& m_{p}(D) \cdot m_{p\in C_i \cap \mathcal{C}}C_i  \\
		%&=& m_{p}(D)\cdot b_i.
		%\end{eqnarray*}
%for each $i$. Summing over all $i$, we get	
		\begin{eqnarray*}
		 D\cdot (C_{1}+ \ldots + C_{d}) &\geq& \sum_{p \in \mathcal{P}}m_{p}(D)\cdot m_{p}(C_{1}+ \ldots + C_{d})\\
		&\stackrel{(*)}{\geq}& 2\sum_{p \in \mathcal{P}} m_{p}(D) \\
		&\stackrel{(\triangle)}{>}& \frac{2(L \cdot D)\sum b_i}{\sum L\cdot C_i},
		\end{eqnarray*}
		where $(*)$ comes from the fact that all the singular points of $C_{1}+ \ldots + C_{d}$ are at least double points. Now we use the nefness 
		of $\frac{d\sum b_i}{(d-1)^2(\sum L\cdot C_i)}L-C_i$ and note the following  inequality: 
		\[
		\frac{d\sum b_i}{(d-1)^2(\sum L\cdot C_i)}L\cdot D \geq D\cdot C_i.
		\] 
		After adding over all $i=1,...,d$, this leads to the following, 
		\begin{eqnarray*}
			\frac{d^2\sum b_i}{(d-1)^2(\sum L\cdot C_i)}L\cdot D &\geq&  D\cdot (C_{1}+ \ldots + C_{d}) > \frac{2(L \cdot D)\sum b_i}{\sum L\cdot C_i} \\
			%\Rightarrow \frac{d^2C_1^2}{d-1} &>& 2(d-1)C_1^2 \quad \quad \text{(since $L$ is ample)} \\
			\Rightarrow \left(\frac{d}{d-1} \right)^2 &>& 2 \quad \quad \text{(since $\sum b_i > 0$ and $L$ is ample)}.
		\end{eqnarray*}
		This is a contradiction, since $d\geq 4$.
	\end{proof}

@@

For an arbitrary curve arrangement $\mathcal{C}$ satisfying Assumption \ref{star1}, the base constant is at most 
$C_1^2(d-1)$. So the following question arises naturally.

@@

 \begin{proof}
Under the given hypotheses, we have $b_i = C_1^2(d-1)$  and $L\cdot C_i = L\cdot L_1$
for all $i$. 
The proof is now immediate from Theorem \ref{main-theorem}.    
 \end{proof} This immediately gives the following.

@@

The proof of Theorem \ref{main-theorem} can be modified to show that for arrangements
satisfying Assumption \ref{star1} which are not star arrangements, we still have the following lower and upper bounds for the Seshadri constants.

@@

Indeed, suppose that the above inequality does not hold. 
As in the proof of Theoerem \ref{main-theorem}, B\'ezout's theorem applied to $D$ and $C_1+\dots+C_d$ gives

@@

\begin{example}
Consider the Hesse arrangement $\mathcal{C} = \{C_1, \ldots, C_{12}\}$  of conics in $\mathbb{P}^2$
which has 21 singular points.  The base constant of $\mathcal{C}$  is 8.  
 The Seshadri constant of $L = \mathcal{O}_{\mathbb{P}^2}(1)$ at these 21 points is known to be $\frac{1}{5}$; see \cite[Example 3.5]{JP}.

@@

We end with an example of a configuration of lines on a K3 surface in $\mathbb{P}^3$. In this example, the nefness hypothesis of Theorem \ref{main-theorem} is not satisfied, but the conclusions hold. 
%Though it does
%not fit in the situation studied in this paper (since the curves in the arrangement are not
%linearly equivalent to each other), it gives an interesting computation of a multi-point Seshadri
%constant on a K3 surface.

@@

Note that Theorem \ref{main-theorem} predicts that the Seshadri constant is indeed $\frac{1}{10}$, though that theorem cannot be applied here. This is because 
$d=48, b_1=\dots = b_{48} = 10$ and $L\cdot l_i = 1$ for every $i$. The divisor
in question is not nef, as can be seen by intersecting with $l_i$.

@@

 \subsection*{Acknowledgements} We thank the referee for a careful reading of the paper, pointing out some corrections and giving several useful comments which substantially improved the paper. The first author is partially supported by a grant from Infosys Foundation.

>>>>

Finally, we give an example (Example \ref{K3}) of a computation of the multi-point Seshadri constant at the singular locus of an arrangement of curves on a K3 surface. In this example, the nefness hypothesis of Theorem \ref{main-theorem} is not true but the conclusion about the Seshadri constant still holds.

@@

  For each $1\le i \le d$, let $b_i =  \# \, C_{i} \cap {\rm Sing}(\mathcal{C})$. Since $\mathcal{C}$ is connected, $b_i >0$ for all $i$.
   We define the \emph{base constant} of $\mathcal{C}$ as 
   $$\text{bs}(\mathcal{C}) = \max\{b_1,\dots, b_d\},$$ i.e., 
$\text{bs}(\mathcal{C})$		
  is equal to the maximum number of singular points  of $\mathcal{C}$ that are contained in a single curve $C_{i} \in \mathcal{C}$.
	%\end{definition}

@@

By \cite[Example 3.5]{JP}, the above question has a negative answer for the Hesse arrangement of conics 
in $\mathbb{P}^2$; see Example \ref{hesse}. However it is open for line arrangements in $\mathbb{P}^2$. 
It is interesting to study the situations in which it is true on arbitrary surfaces. Our main theorem below answers this question positively under an additional nefness condition.

@@

	 Suppose that there exists an irreducible and reduced curve $D\notin \mathcal{C}$ having multiplicity $m_{p}(D)$ at each point $p \in \mathcal{P}$ such that $D \cap \mathcal{P} \ne \emptyset$ and 
		%$$\frac{L \cdot D}{\sum_{p \in \mathcal{P}} m_{p}(D)} < \frac{L\cdot C_1}{C_1^2(d-1)}.$$
		$$\frac{L \cdot D}{\sum_{p \in \mathcal{P}} m_{p}(D)} < \min_{1\le i \le d}\left\{\frac{L\cdot C_i}{b_i}\right\}.$$
So for each $1 \le i \le d$, we have 
                $$b_i(L \cdot D)  < (L\cdot C_i) \sum_{p \in \mathcal{P}} m_{p}(D).$$

Adding over all $i$, we have 
		$$(\triangle) : \quad (L \cdot D)\sum_{i=1}^d b_i < \left(\sum_{p \in \mathcal{P}} m_{p}(D)\right)\sum_{i=1}^d L\cdot C_i.$$
		By B\'ezout's theorem applied to $D$ and $C_1+\dots+C_d$, we obtain
		%\begin{eqnarray*}
		%D\cdot C_i &\geq& m_{p}(D) \cdot m_{p\in C_i \cap \mathcal{C}}C_i  \\
		%&=& m_{p}(D)\cdot b_i.
		%\end{eqnarray*}
%for each $i$. Summing over all $i$, we get	
		\begin{eqnarray*}
		 D\cdot (C_{1}+ \ldots + C_{d}) &\geq& \sum_{p \in \mathcal{P}}m_{p}(D)\cdot m_{p}(C_{1}+ \ldots + C_{d})\\
		&\stackrel{(*)}{\geq}& 2\sum_{p \in \mathcal{P}} m_{p}(D) \\
		&\stackrel{(\triangle)}{>}& \frac{2(L \cdot D)\sum b_i}{\sum L\cdot C_i},
		\end{eqnarray*}
		where $(*)$ comes from the fact that all the singular points of $C_{1}+ \ldots + C_{d}$ are at least double points. Now we use the nefness 
		of $\frac{d\sum b_i}{(d-1)^2(\sum L\cdot C_i)}L-C_i$ and note the following  inequality: 
		\[
		\frac{d\sum b_i}{(d-1)^2(\sum L\cdot C_i)}L\cdot D \geq D\cdot C_i.
		\] 
		After adding over all $i=1,...,d$, this leads to the following, 
		\begin{eqnarray*}
			\frac{d^2\sum b_i}{(d-1)^2(\sum L\cdot C_i)}L\cdot D &\geq&  D\cdot (C_{1}+ \ldots + C_{d}) > \frac{2(L \cdot D)\sum b_i}{\sum L\cdot C_i} \\
			%\Rightarrow \frac{d^2C_1^2}{d-1} &>& 2(d-1)C_1^2 \quad \quad \text{(since $L$ is ample)} \\
			\Rightarrow \left(\frac{d}{d-1} \right)^2 &>& 2 \quad \quad \text{(since $\sum b_i > 0$ and $L$ is ample)}.
		\end{eqnarray*}
		This is a contradiction, since $d\geq 4$.
	\end{proof}

@@

For an arbitrary curve arrangement $\mathcal{C}$ satisfying Assumption \ref{star1}, the base constant is at most 
$C_1^2(d-1)$. So the following question arises naturally from
Question \ref{question}.

@@

 \begin{proof}
Under the given hypotheses, we have $b_i = C_1^2(d-1)$  and $L\cdot C_i = L\cdot C_1$
for all $i$. 
The proof is now immediate from Theorem \ref{main-theorem}.    
 \end{proof} This immediately gives the following.

@@

The proof of Theorem \ref{main-theorem} can be modified to show that for arrangements
satisfying Assumption \ref{star1} which are not necessarily star arrangements, we still have the following lower and upper bounds for the Seshadri constants.

@@

Indeed, suppose that the above inequality does not hold. 
As in the proof of Theorem \ref{main-theorem}, B\'ezout's theorem applied to $D$ and $C_1+\dots+C_d$ gives

@@

\begin{example}\label{hesse}
Consider the Hesse arrangement $\mathcal{C} = \{C_1, \ldots, C_{12}\}$  of conics in $\mathbb{P}^2$
which has 21 singular points.  The base constant of $\mathcal{C}$  is 8.  
 The Seshadri constant of $L = \mathcal{O}_{\mathbb{P}^2}(1)$ at these 21 points is known to be $\frac{1}{5}$; see \cite[Example 3.5]{JP}.

@@

We end with an example of a configuration of lines on a K3 surface in $\mathbb{P}^3$. In this example, the nefness hypothesis of Theorem \ref{main-theorem} is not satisfied, but the conclusion holds. 
%Though it does
%not fit in the situation studied in this paper (since the curves in the arrangement are not
%linearly equivalent to each other), it gives an interesting computation of a multi-point Seshadri
%constant on a K3 surface.

@@

Note that Theorem \ref{main-theorem} predicts that the Seshadri constant is indeed $\frac{1}{10}$, though  it cannot be applied here. This is because 
$d=48, b_i = 10$ and $L\cdot l_i = 1$ for every $i$. So the divisor which appears in the statement of Theorem \ref{main-theorem}
is not nef, as can be seen by intersecting with $l_i$.

@@

 \subsection*{Acknowledgements} We thank the referee for a careful reading of the paper, pointing out some corrections and giving several useful suggestions which substantially improved the paper. The first author is partially supported by a grant from Infosys Foundation.
